---
title: "Stat 437 HW2"
author: 
     - Yu-Tung (Jenny), Cheng (ID:11678647)
header-includes:
   - \usepackage{bbm}
   - \usepackage{amssymb}
   - \usepackage{amsmath}
   - \usepackage{graphicx,float}
   - \usepackage{natbib}
output:
  html_document: default
fontsize: 11pt
---

```{r, echo=FALSE, warning=FALSE}
library(knitr)
opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# General rule

Please show your work and submit your computer codes in order to get points. Providing correct answers without supporting details does not receive full credits. This HW covers:

- Advanced Visualizations via ggplot2: adjusting legends, fonts, orientation, and math expressions
- Visualizing networks as graphs
- Interactive visualization 

For an assignment or project, you DO NOT have to submit your answers or reports using typesetting software. However, your answers must be well organized and well legible for grading. Please upload your answers in a document to the course space. Specifically, if you are not able to knit a .Rmd/.rmd file into an output file such as a .pdf, .doc, .docx or .html file that contains your codes, outputs from your codes, your interpretations on the outputs, and your answers in text (possibly with math expressions), please organize your codes, their outputs and your answers in a document in the format given below:

```
Problem or task or question ... 
Codes ...
Outputs ...
Your interpretations ...
```

It is absolutely not OK to just submit your codes only. This will result in a considerable loss of points on your assignments or projects. 

# Problem 1

Please refer to the NYC flight data `nycflights13` that has been discussed in the lecture notes and whose manual can be found at https://cran.r-project.org/web/packages/nycflights13/index.html. We will use `flights`, a tibble from `nycflights13`.


You are interested in looking into the average `arr_delay` for 4 different `month` 12, 1, 7 and 8, for 3 different `carrier` "UA", "AA" and "DL", and for `distance` that are greater than 700 miles, since you suspect that colder months and longer distances may result in longer average arrival delays. Note that you need to extract observations from `flights`, and that you are required to use `dplyr` for this purpose. 

The following tasks and questions are based on the extracted observations.
```{r, cache=TRUE}
library(dplyr)
library(nycflights13)
library(ggplot2)
library(tidyverse)
head(flights)

##create a small data frame 'temp' from flights and select 6 months and 3 carriers from 'flights'

temp = flights %>% select(month, arr_delay,carrier, distance) %>% filter(month %in% c(12,1,7,8), carrier %in% c("UA", "AA", "DL"), distance > 700)

#remove the missing value NA
temp = na.omit(temp)
myData = temp %>% group_by(carrier, month) %>% summarise_at(vars(arr_delay, distance), list(mean))

temp

```

(1.a) For each combination of the values of `carrier` and `month`, obtain the average `arr_delay` and obtain the average `distance`. Plot the average `arr_delay` against the average `distance`, use `carrier` as facet; add a title "Base plot" and center the title in the plot. This will be your base plot, say, as object `p`. Show the plot `p`.
```{r, cache=TRUE}

p = ggplot(data = myData, aes(y= arr_delay, x = distance))+geom_point(col="blue")

p + facet_wrap(~carrier, nrow = 1)+labs(x= 'average_distance', y='average_arr_delay',title=('Base Plot'))+theme(plot.title= element_text(hjust = 0.5), )


##Interpretation: In this Base Plot, it indicates that UA focuses on long-distance flights compared to AA and DL. DL relatively shows the higher frequency of the average of arrival delay in the short-distance flights compared with AA.  

```

(1.b) Modify `p` as follows to get a plot `p1`: connect the points for each `carrier` via one type of dashed line; code the 3 levels of `carrier` as $\alpha_1$, $\beta_{1,2}$ and $\gamma^{[0]}$, and display them in the strip texts; change the legend title into "My $\zeta$" (this legend is induced when you connect points for each `carrier` by a type of line), and put the legend in horizontal direction at the bottom of the plot; add a title "With math expressions" and center the title in the plot. Show the plot `p1`.

```{r, cache=TRUE}

#map 3 levels to expressions
carrierStg = c(expression(alpha[1]), expression(beta['1,2']), expression(gamma^{'[0]'}))

#create variable DF (a factor) with levels “α1”, “β2” and “gamma”
myData$DF = factor(myData$carrier, labels = carrierStg) 

#use 'slice' to check correctness of mapping

myData %>% select(distance,arr_delay,DF, carrier) %>% group_by(carrier) %>% slice(1)


p1 = ggplot(myData, aes(distance, arr_delay))+geom_point(color = "steelblue")+ theme(plot.title = element_text(hjust=0.5))+geom_line(aes(linetype= myData$DF, color = "red"))+labs(linetype = expression(paste("My ", zeta, sep=""))) + scale_linetype_discrete(labels =carrierStg) 


p1 = p1+facet_wrap(~DF, nrow = 1, labeller = label_parsed) + labs(x='ave_distance', y='ave_arr_delay', title=('Base Plot \n With Math Expression'))+theme(legend.position="bottom",legend.direction="horizontal")

print(p1)

##Interpretation: The graph shows the changes of the 3 levels of carriers to the Mathematical expressions which are AA as α1 with the solid line to connect the data points,  DL as β1,2 with the dash line to connect the data points, and UA as γ^[0] with looser dash line to connect the data points. From the visualization, DL as β1,2 displays relatively huge gap between their less frequent average of arrival delay and their most frequent data point. By this case, among the AA, DL and UA, DL(β1,2) has comparatively highly unstable for the average flight punctuality rate.  

```

(1.c) Modify `p1` as follows to get a plot `p2`: set the font size of strip text to be 12 and rotate the strip texts counterclockwise by 15 degrees; set the font size of the x-axis text to be 10 and rotate the x-axis text clockwise by 30 degrees; set the x-axis label as "$\hat{\mu}$ for mean arrival delay"; add a title "With front and text adjustments" and center the title in the plot. Show the plot `p2`

```{r, cache=TRUE}

p2 = p1+ theme(axis.text.x = element_text(size= 10, angle = -30), axis.title.x= element_text(size=10, angle=15), strip.text= element_text(size=12, angle= 15))+labs(x = expression(paste(hat(mu), ' \t mean of arrival delay')), title= ('Base Plot\n With Math Expression \n With Front and Next Adjustment'))

print(p2)

##Interpretation: We set the font size of the strip text of α1, β1,2 and  γ^[0] with the counterclockwise rotate by 15 degrees. Besides that, by adding the mu^hat expression to express the mean of arrival delay with the font size of 10 and be rotated clockwise by 30 degrees. 

```

# Problem 2 

This problem requires you to visualize the binary relationship between members of a karate club as an undirected graph. Please install the R library `igraphdata`, from which you can obtain the data set `karate` and work on it. Create a graph for `karate`. Once you obtain the graph, you will see that each vertex is annotated by a number or letter. What do the numbers or letters refer to? Do you see subgraphs of the graph? If so, what do these subgraphs mean? 

```{r, cache=TRUE}

#install package 'igraphdata' and 'igraph'
library(igraphdata)
library(igraph)
data(karate)
plot(karate)

##Interpretation: This is a graphical representation of the social relationships among 34 individuals in a karate club. Each sub-bubbles of numbers represent to the members in the karate club, and one of the two main bubbles 'A' refers to the karate president John A. and another one with 'H' refers to the karate instructor.  The vertex attribute is regarded as the participant's membership. However,  Actor 9 has more connections to 'H's group. Besides that, some sub-bubble individuals between 'H' and 'A' with the middle part usually show more complexity of overlapping relationships between them. 
```

# Problem 3

This problem requires to to create an interactive plot using `plotly`. If you want to display properly the plot in your HW answers, you may well need to set your HW document as an html file (instead of doc, docx or pdf file) when you compile your R codes.

Please use the `mpg` data set we have discussed in the lectures. Create an interactive, scatter plot between "highway miles per gallon" `hwy` (on the y-axis) and "engine displacement in litres" `displ` (on the x-axis) with the `color` aesthetic designated by "number of cylinders" `cyl`, and set the x-axis label as "engine displacement in litres" and y-axis label as "highway miles per gallon". You need to check the object type for `cyl` and set it correctly when creating the plot. Add the title "# of cylinders" to the legend and adjust the vertical position of the legend, if you can. For the last, you may look through https://plotly.com/r/legend/ for help.

```{r, cache=TRUE}
library(plotly)
library(ggplot2)

plot_ly(data= mpg,x = ~displ, y= ~hwy, color = ~cyl, type = "scatter", symbol = ~cyl, width = 700, height = 400, size = 3, alpha = 0.3, text = ~cyl) %>%
      layout(xaxis = list(title="Engine Displacement in liters"), yaxis = list(title='Highway Miles per gallon'),
             legend=list(title=list(text='# of cylinders')))

##Interpretation: As the visualization shows that set of 4 cylinders has the least engine displacement in liters but it has the largest amount consumption in highway miles per gallon compared to others with 5, 6 and 8 cylinders. However, the cars with 8 cylinders, which are the indicated as round shape of data points, have the pretty high frequency of engine displacement in liters with unstable ranges from different data. The 5 cylinder only has two data points from the graph in the triangular shape of data points with technically fewer sample observations among others.
```